---
title: "pm566 midterm"
author: "Changqing Su"
date: "2020/10/9"
output: html_document

---

## Intoduction 

At the end of 2019, a novel coronavirus was identified as the cause of a cluster of pneumonia cases in China. It rapidly spread, resulting in an epidemic throughout the world. In February 2020, the World Health Organization designated the disease COVID-19, which stands for coronavirus disease 2019.
This report is designed to explore the association beween Covid-19 death and state, age groups.



## Methods

The data was obstained from the US CDC, Centers of Disease Control and Prevention. The date include deaths involving coronavirus disease 2019 (COVID-19), pneumonia, and influenza reported to NCHS by sex and age group and state.

We calculate the proportionate mortality ratio due to COVID-19 by using the COVID-19 deaths/total deaths.

```{r, warning=FALSE, message=FALSE}
library(data.table)
library(dtplyr)
library(dplyr)
library(leaflet)
library(tidyverse)
library(ggplot2)

```



```{r}

covid = fread("data/Provisional_COVID-19_Death_Counts_by_Sex__Age__and_State.csv")

#calculate proportionate mortality ratio due to COVID-19
covid$PMR=covid$`COVID-19 Deaths`/covid$`Total Deaths`

```



#### Compare the proportionate mortality ratio among states.

Here, we create a map of proportionate mortality ratio among the states.



```{r, echo=T, warning=FALSE, message=FALSE, results='hide'}
library(tigris)
states <- states(cb=T)
```



```{r, warning=F, message=F}
covid_state=covid[which(covid$Sex == "All Sexes" & covid$State != "United States" & covid$`Age group`=="All Ages"),]


#covid_state$State[which(covid_state$State %in% states$NAME ==F)]


covid_state[33,7:12]=covid_state[33,7:12]+covid_state[34,7:12]
covid_state=covid_state[-34, ]
covid_state$PMR=covid_state$`COVID-19 Deaths`/covid_state$`Total Deaths`
colnames(covid_state)[4]="NAME"


mergedata=merge(x = states, y = covid_state, by = "NAME", all.x = TRUE)


pal <- colorBin("YlOrRd", domain = mergedata$PMR)


leaflet() %>%
 addProviderTiles("CartoDB.Positron") %>%
  setView(-98.483330, 38.712046, zoom = 4) %>% 
  addPolygons(
    data=mergedata,
    fillColor = ~pal(mergedata$PMR),
    fillOpacity = 0.7, 
              weight = 0.2, 
              smoothFactor = 0.2 
  )%>%
  addLegend(pal = pal, 
            values = mergedata$PMR, 
            position = "bottomright", 
            title = "PMR")
  


```

Here we can see that the New York State has the highest proportionate mortality ratio due to COVID-19.
It means that New York City has the most people died because of Covid-19.


```{r, warning=F, message=F}
pal2 <- colorBin("YlOrRd", domain = mergedata$`COVID-19 Deaths`)

leaflet() %>%
 addProviderTiles("CartoDB.Positron") %>%
  setView(-98.483330, 38.712046, zoom = 4) %>% 
  addPolygons(
    data=mergedata,
    fillColor = ~pal2(mergedata$`COVID-19 Deaths`),
    fillOpacity = 0.7, 
              weight = 0.2, 
              smoothFactor = 0.2 
  )%>%
  addLegend(pal = pal2, 
            values = mergedata$`COVID-19 Deaths`, 
            position = "bottomright", 
            title = "Total Deaths")
  

```
According to this map, we can see that the New York State has the most Covid-19 deaths(more than 30,000).
California has the second most Covid-19 deaths(more than 15,000). But comparing to the previous map, it does not have a high RMR. This means that California has a large number because of the large population.





#### Compare the proportionate mortality ratio beween genders and age groups.



```{r , warning=F, message=F}
covid_2=covid[which(covid$Sex == "All Sexes" & covid$State == "United States" & covid$`Age group`!="All Ages"),]
covid_2=covid_2[which(covid_2$`Age group` == "Under 1 year" | covid$`Age group` == "1-4 years" | covid$`Age group` == "5-14 years" | covid$`Age group` == "15-24 years" | covid$`Age group` == "25-34 years" | covid$`Age group` == "35-44 years" | covid$`Age group` == "45-54 years" | covid$`Age group` == "55-64 years" | covid$`Age group` == "65-74 years" | covid$`Age group` == "75-84 years" | covid$`Age group` == "85 years and over")]

covid_2=na.omit(covid_2)

covid_3=subset(covid_2, select = c("Sex", "Age group","COVID-19 Deaths","Total Deaths","PMR"))
                
knitr::kable(covid_3) 
```

## Conlucsion
From the map we can conlucde that the New York State has the highest proportionate mortality ratio due to COVID-19.
We found that older people (greater than 65 years old) are easier die because of Covid-19.
